---
title: "AQ_Cell_counts"
output: html_document
date: "2025-03-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly) # Turns visualizations interactive
library(moderndive)
library(readxl)
#library(ggpubr)
library(RColorBrewer)
library(devtools)
library(Routliers)
library(htmlwidgets)

#knitr::opts_knit$set(root.dir = normalizePath("C:/Users/antho/OneDrive - University of Massachusetts/Richardson Lab/Anthony R Files/Data Files"))

# For Laptop
setwd("C:/Users/Thinkpad-X1/OneDrive - University of Massachusetts/Richardson Lab/Anthony R Files/Data Files")


```

Importing data

```{r}
library(readr)
# PC
cell_counts <- read_csv("C:/Users/antho/OneDrive - University of Massachusetts/Richardson Lab/Anthony R Files/Data Files/CREST20X_data_6.csv")

# Lab PC
cell_counts <- read_csv("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files/CREST20X_data_6.csv")

# Laptop
cell_counts <- read_csv("C:/Users/Thinkpad-X1/OneDrive - University of Massachusetts/Richardson Lab/Anthony R Files/Data Files/CREST20X_data_6.csv")

```

(FIRST STEP) Averaging cell density between left and right hemispheres

```{r}

# Avg between hemispheres for DAPI, CC1, ASPA density
cell_counts %>%
  group_by(Season, `Bregma(mm)`, ROI, MouseID) %>%
  summarize(mean(TotalDAPIDensitymm2), mean(TotalCC1Densitymm2), mean(TotalASPADensitymm2)) -> avg_density

# Renaming the columns for cleanliness
avg_density %>%
  rename(DAPIdensity = `mean(TotalDAPIDensitymm2)`) %>%
  rename(CC1density = `mean(TotalCC1Densitymm2)`) %>%
  rename(ASPAdensity = `mean(TotalASPADensitymm2)`) -> avg_density

```


Manipulate tibble to longer format

```{r}
long_data <- avg_density %>%
  pivot_longer(cols = c(DAPIdensity, CC1density, ASPAdensity), # Adjust these if different
               names_to = "Cell_Type",
               values_to = "Densities")

#write.csv(long_data, file = "cellCountslongData.csv")
```


Splitting data between ROIs

```{r}
cell_counts %>%
   filter(ROI == "CC") -> CC_group

cell_counts %>%
   filter(ROI == "Cg") -> Cg_group
```


(FIXME) Making function to identify outliers in data

```{r}
cell_counts %>%
   select(ROI,`Total DAPI`, bregma_group) %>%
   group_by(ROI, bregma_group) %>%
   summarize(Q1 = quantile(`Total DAPI`, 0.25),
             Q3 = quantile(`Total DAPI`, 0.75),
             IQR = Q3 - Q1,
             lower_bound = Q1 - 1.5 * IQR,
             upper_bound = Q3 + 1.5 * IQR)
      






find_outlier <- function(df, values_col, groups_col) {
   df %>%
      group_by(groups_col) %>%
      summarize(Q1 = quantile(value_col, 0.25, na.rm = TRUE), 
             Q3 = quantile(value_col, 0.75, na.rm = TRUE),
             IQR = Q3 - Q1,
             lower_bound = Q1 - 1.5 * IQR,
             upper_bound = Q3 + 1.5 * IQR,
             is_outlier = ifelse((value_col < lower_bound | value_col > upper_bound), 1, 0)) %>%
      filter(is_outlier == 1)
}



```


Plotting Data without removing outliers

```{r}
cell_counts %>%
   mutate(`Bregma(mm)` = as.factor(`Bregma(mm)`)) -> cell_counts # Turns Bregma into a categorical variable

cell_counts %>%
   mutate(bregma_group = interaction(`Bregma(mm)`, Group, sep = "mm | ")) -> cell_counts


plot <- ggplot(data = cell_counts, aes(y = TotalDAPIDensitymm2, x = bregma_group, fill = bregma_group)) +
   facet_wrap("ROI") +
   geom_boxplot() +
   theme_minimal(base_size = 12) +  # Modern, clean theme with larger font
   theme(plot.title = element_text(hjust = 0.5),  
        legend.position = "top") +
  labs(title = "Total DAPI Density (mm^2) by Group",
       x = "Bregma (mm)",
       y = "Total DAPI Density (mm^2)") +
   theme(panel.grid.major = element_line(color = "grey80", linetype = "dashed"),
        panel.grid.minor = element_blank(),
        legend.position = "top",
        strip.background = element_rect(fill = "lightgrey", color = "black")) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
         strip.text = element_text(size = 14),
         legend.position = "top") +
   theme(plot.margin = margin(10, 10, 40, 10)) +  # Increase bottom margin
   theme(axis.title.x = element_text(margin = margin(t = 30))) 

   
ggplotly(plot)  


saveWidget(ggplotly(plot), "interactive_plot2.html", selfcontained = TRUE)

```


Same thing but separated by Group
(cc group)
```{r}
cell_counts %>%
   mutate(bregma_group = interaction(`Bregma(mm)`, Group, sep = "mm | ")) -> cell_counts
          
cc_plot <- ggplot(data = CC_group, aes(y = TotalDAPIDensitymm2, x = bregma_group, fill= bregma_group)) +
   #facet_wrap("ROI") +
   geom_boxplot() +
   theme_minimal(base_size = 12) +  # Modern, clean theme with larger font
   theme(plot.title = element_text(hjust = 0.5),  
        legend.position = "top") +
  labs(title = "DAPI Density (mm^2) (CC Group)",
       x = "Bregma (mm)",
       y = "Total DAPI") +
   theme(panel.grid.major = element_line(color = "grey80", linetype = "dashed"),
        panel.grid.minor = element_blank(),
        legend.position = "top",
        strip.background = element_rect(fill = "lightgrey", color = "black")) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
         strip.text = element_text(size = 14),
         legend.position = "top") +
   theme(plot.margin = margin(10, 10, 40, 10)) +  # Increase bottom margin
   theme(axis.title.x = element_text(margin = margin(t = 20))) 

   
   
ggplotly(cc_plot)  
saveWidget(ggplotly(cc_plot), "interactive_plot.html", selfcontained = TRUE)

```


Same thing but separated by Group
(cg group)
```{r}
cg_plot <- ggplot(data = Cg_Group, aes(y = `Total DAPI`, x = bregma_group, fill= bregma_group)) +
   #facet_wrap("ROI") +
   geom_boxplot() +
   theme_minimal(base_size = 12) +  # Modern, clean theme with larger font
   theme(plot.title = element_text(hjust = 0.5),  
        legend.position = "top") +
  labs(title = "Cell Count (Cg Group)",
       x = "Bregma (mm)",
       y = "Total DAPI") +
   theme(panel.grid.major = element_line(color = "grey80", linetype = "dashed"),
        panel.grid.minor = element_blank(),
        legend.position = "top",
        strip.background = element_rect(fill = "lightgrey", color = "black")) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
         strip.text = element_text(size = 14),
         legend.position = "top") +
   theme(plot.margin = margin(10, 10, 40, 10)) +  # Increase bottom margin
   theme(axis.title.x = element_text(margin = margin(t = 20))) 

   
   
ggplotly(cg_plot)  
```


Checking spreadsheet

```{r}
max(cell_counts$TotalDAPIDensitymm2)

# THIS IS THE MISSING STEP
# Avg between hemispheres for DAPI density
cell_counts %>%
  group_by(Group, `Bregma(mm)`, ROI, MouseID) %>%
  summarize(mean(TotalDAPIDensitymm2)) -> avg_hemisphere

# Avg between hemispheres for CC1 density
cell_counts %>%
  group_by(Group, `Bregma(mm)`, ROI, MouseID) %>%
  summarize(mean(TotalCC1Densitymm2)) -> avg_hemisphere


```


NEW PLOT (After calculating average DAPI Density between the 2 hemispheres)

```{r}
avg_hemisphere %>%
   mutate(bregma_group = interaction(`Bregma(mm)`, Group, sep = "mm | ")) -> avg_hemisphere

plot_avg <- ggplot(data = avg_hemisphere, aes(y = `mean(TotalCC1Densitymm2)`, x = bregma_group, fill = bregma_group)) +
   facet_wrap("ROI") +
   geom_boxplot() +
   theme_minimal(base_size = 12) +  # Modern, clean theme with larger font
   theme(plot.title = element_text(hjust = 0.5),  
        legend.position = "top") +
  labs(title = "Total CC1 Density (mm^2) by Group",
       x = "Bregma (mm) and Treatment",
       y = "Total CC1 Density (mm^2)") +
   theme(panel.grid.major = element_line(color = "grey80", linetype = "dashed"),
        panel.grid.minor = element_blank(),
        legend.position = "top",
        strip.background = element_rect(fill = "lightgrey", color = "black")) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
         strip.text = element_text(size = 14),
         legend.position = "top") +
   theme(plot.margin = margin(10, 10, 40, 10)) +  # Increase bottom margin
   theme(axis.title.x = element_text(margin = margin(t = 30))) 

   
ggplotly(plot_avg) 

saveWidget(ggplotly(plot_avg), "plot_avg.html", selfcontained = TRUE)


```



Splitting between Seasons

```{r}
avg_density %>%
  filter(Season == "Fall") -> fall

long_data %>%
  filter(Season == "Spring") -> spring

long_data %>%
  filter(Season == "Summer") -> summer

long_data %>%
  filter(Season == "Winter") -> winter

```


(OLD) Trying Routliers package 

```{r}

# DAPI Outliers per Season
outliers_mad(x = fall$DAPIdensity, threshold = 2) -> DAPIfall_outlier_info
outliers_mad(x = spring$DAPIdensity, threshold = 2) -> DAPIspring_outlier_info
outliers_mad(x = summer$DAPIdensity, threshold = 2) -> DAPIsummer_outlier_info
outliers_mad(x = winter$DAPIdensity, threshold = 2) -> DAPIwinter_outlier_info
# Threshold argument controls number of MAD plus or minus the median to consider an outlier

# CC1 Outliers per Season
outliers_mad(x = fall$CC1density, threshold = 2) -> CC1fall_outlier_info
outliers_mad(x = spring$CC1density, threshold = 2) -> CC1spring_outlier_info
outliers_mad(x = summer$CC1density, threshold = 2) -> CC1summer_outlier_info
outliers_mad(x = winter$CC1density, threshold = 2) -> CC1winter_outlier_info

# ASPA Outliers Per Season
outliers_mad(x = fall$ASPAdensity, threshold = 2) -> ASPAfall_outlier_info
outliers_mad(x = spring$ASPAdensity, threshold = 2) -> ASPAspring_outlier_info
outliers_mad(x = summer$ASPAdensity, threshold = 2) -> ASPAsummer_outlier_info
outliers_mad(x = winter$ASPAdensity, threshold = 2) -> ASPAwinter_outlier_info

# To extract the upper and lower bounds in variable form
avg_density %>% 
  mutate(DAPIoutlier = ifelse(
    (DAPI_outlier_info$LL_CI_MAD <= AvgDAPIdensity & DAPI_outlier_info$UL_CI_MAD >= AvgDAPIdensity), 0, 1)) %>%
  mutate(CC1outlier = ifelse(
    (CC1_outlier_info$LL_CI_MAD <= AvgCC1density & CC1_outlier_info$UL_CI_MAD >= AvgCC1density), 0, 1)) %>%
  mutate(ASPAoutlier = ifelse(
    (ASPA_outlier_info$LL_CI_MAD <= AvgASPAdensity & ASPA_outlier_info$UL_CI_MAD >= AvgASPAdensity), 0, 1))-> avg_density
```

Function to automate finding outlier info

```{r}
calculate_outliers <- function(df) {
  if (nrow(df) < 3) {  # Ensure at least 3 values for meaningful MAD calculation
    df <- df %>%
      mutate(Lower_Bound = NA, Upper_Bound = NA, Outlier = FALSE)
  } else {
    bounds <- outliers_mad(df$Densities)  # Calculate MAD-based outlier bounds
    df <- df %>%
      mutate(Median = bounds$Median,
             Lower_Bound = bounds$LL_CI_MAD, 
             Upper_Bound = bounds$UL_CI_MAD,
             Outlier = Densities < Lower_Bound | Densities > Upper_Bound) %>% # Flag outliers
      mutate(Outlier = ifelse(Outlier == TRUE, 1, 0))
    
  }
  return(df)
}

long_data %>%
  group_by(Season, `Bregma(mm)`, Cell_Type, ROI) %>%
  group_modify(~ calculate_outliers(.x)) %>%
  ungroup() -> all_outlier_info

# Checking correctness
avg_density %>%
  filter(Season == "Fall") %>%
  filter(`Bregma(mm)` == 0.9) %>%
  filter(ROI == "CC") %>%
  filter(MouseID == "M4613")


```



Visualizing number of outliers +/- 2 MAD from the median value

```{r}
dapi <- outliers_mad(x = avg_density$AvgDAPIdensity, threshold = 2)
plot_outliers_mad(dapi, x = avg_density$AvgDAPIdensity)

cc1 <- outliers_mad(x = avg_density$AvgCC1density, threshold = 2)
plot_outliers_mad(cc1, x = avg_density$AvgDAPIdensity)

aspa <- outliers_mad(x = avg_density$AvgASPAdensity, threshold = 2)
plot_outliers_mad(aspa, x = avg_density$AvgDAPIdensity)
```



Creating an interaction column between Season and Bregma

```{r}
avg_density %>%
   mutate(bregma_group = interaction(`Bregma(mm)`, Season, sep = "mm | ")) -> avg_density

long_data %>%
   mutate(bregma_group = interaction(`Bregma(mm)`, Season, sep = "mm | ")) -> long_data

all_outlier_info %>%
   mutate(bregma_group = interaction(`Bregma(mm)`, Season, sep = "mm | ")) -> all_outlier_info
```

(DAPI)
Starting cell counts plot INCLUDING geom jitter for outliers

```{r}
plot_avg_DAPI <- ggplot(data = (all_outlier_info%>% filter(Cell_Type == "DAPIdensity")),
                        aes(y = Densities,
                            x = bregma_group,
                            fill = bregma_group)) +
  # Plotting data spread
  geom_boxplot(outliers = FALSE) +
  
  # Plotting non-outliers
  geom_jitter(data = (all_outlier_info %>%
                        filter(Cell_Type == "DAPIdensity") %>%
                        filter(Outlier == "0")),
              color = "black",
              shape = 1) + 
  facet_wrap("ROI") +
  
  # Plotting outliers
  geom_jitter(data = all_outlier_info %>%
                filter(Outlier == 1) %>%
                filter(Cell_Type == "DAPIdensity"), color = "red") +
  
  # Labels
  labs(title = "Total DAPI Density (mm^2) by Group",
       x = "Bregma (mm) and Treatment",
       y = "Total DAPI Density (mm^2)") +
  
  # Visual Changes
  scale_fill_grey(start = 0.8, end = 0.2) +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 

  

plot_avg_DAPI

#ggplotly(plot_avg_DAPI)

saveWidget(ggplotly(plot_avg_DAPI), "DAPIplot.html", selfcontained = TRUE)
```
(Useless)
```{r}
p <- ggplotly(plot_avg_DAPI)

# Iterate over each trace and disable the outlier points if the trace has the 'boxpoints' attribute
for (i in seq_along(p$x$data)) {
  if (!is.null(p$x$data[[i]]$boxpoints)) {
    p$x$data[[i]]$boxpoints <- "false"
  }
}

# Display the updated plot
p
```



(CC1)

```{r}
plot_avg_CC1 <- ggplot(data = (all_outlier_info%>% filter(Cell_Type == "CC1density")),
                        aes(y = Densities,
                            x = bregma_group,
                            fill = bregma_group)) +
  geom_boxplot() +
  facet_wrap("ROI") +
 # stat_summary(fun = median, geom = "errorbar") + 
  geom_jitter(data = all_outlier_info %>%
                filter(Outlier == 1) %>%
                filter(Cell_Type == "CC1density")) +
  labs(title = "Total CC1 Density (mm^2) by Group",
       x = "Bregma (mm) and Treatment",
       y = "Total CC1 Density (mm^2)") +
  

  scale_fill_grey(start = 0.8, end = 0.2) +
  theme_minimal(base_size = 12) +  # Modern, clean theme with larger font
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(panel.grid.major = element_line(color = "grey80", linetype = "dashed"),
        panel.grid.minor = element_blank(),
        legend.position = "right",
        strip.background = element_rect(fill = "lightgrey", color = "black"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 14),
        plot.margin = margin(10, 10, 40, 10),
        axis.title.x = element_text(margin = margin(t = 30))) 
  

plot_avg_CC1


#ggplotly(plot_avg_CC1)

saveWidget(ggplotly(plot_avg_CC1), "CC1plot.html", selfcontained = TRUE)
```


(ASPA)

```{r}
plot_avg_ASPA <- ggplot(data = (all_outlier_info%>% filter(Cell_Type == "ASPAdensity")),
                        aes(y = Densities,
                            x = bregma_group,
                            fill = bregma_group)) +
  facet_wrap("ROI") +
  stat_summary(fun = median, geom = "col", width = 0.5) + 
  geom_jitter(data = all_outlier_info %>%
                filter(Outlier == 1) %>%
                filter(Cell_Type == "ASPAdensity")) +
  scale_fill_grey(start = 0.8, end = 0.2) +
  theme_minimal(base_size = 12) +  # Modern, clean theme with larger font
  theme(plot.title = element_text(hjust = 0.5),  
        legend.position = "top") +
  labs(title = "Total ASPA Density (mm^2) by Group",
       x = "Bregma (mm) and Treatment",
       y = "Total ASPA Density (mm^2)") +
  theme(panel.grid.major = element_line(color = "grey80", linetype = "dashed"),
        panel.grid.minor = element_blank(),
        legend.position = "top",
        strip.background = element_rect(fill = "lightgrey", color = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
         strip.text = element_text(size = 14),
         legend.position = "top") +
  theme(plot.margin = margin(10, 10, 40, 10)) +  # Increase bottom margin
  theme(axis.title.x = element_text(margin = margin(t = 30))) 

#plot_avg_DAPI

ggplotly(plot_avg_ASPA)

saveWidget(ggplotly(plot_avg_ASPA), "ASPAplot.html", selfcontained = TRUE)
```




