---
title: "AQ_Correlation_Graph-241104.08"
output: pdf_document
date: "2024-11-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(lubridate)
library(dplyr)
library(dbplot)
library(stringr)
library(tidyr)
library(GGally)
library(skimr)
library(moderndive)
library(readxl)
 
#setwd("C:/Users/lab/OneDrive - University of Massachusetts (2)/Anthony R Files/Data Files")
```

IMPORTING DATA

```{r}
library(readr)

# First 4 day period
liqhd_data1a <- read_csv("24110401.CSV")
liqhd_data1b <- read_csv("24110500.CSV")
liqhd_data1c <- read_csv("24110700.CSV")

# Second 4 day period
liqhd_data2a <- read_csv("24111201.CSV")


# Bottle weights data looks like this because I manipulated it in Excel
bottle_weights <- read_csv("Bottle_weights.csv")

### Importing Data for PC

liqhd_data1a <- read_csv("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files/24110401.CSV", col_select = -...220)
liqhd_data1b <- read_csv("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files/24110500.CSV", col_select = -...220)
liqhd_data1c <- read_csv("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files/24110700.CSV", col_select = -...220)


liqhd_data2a <- read_csv("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files/24111201.CSV")

liqhd_data3a <- read_csv("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files/24111800.CSV")
liqhd_data3b <- read_csv("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files/24112000.CSV")




bottle_weights <- read_csv("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files/Bottle_weights.csv")


# New approach so that adding more data will be easier
# I am combining all liqhd data into one big master dataset as they come in

rbind(liqhd_data1a, liqhd_data1b, liqhd_data1c, liqhd_data2a, liqhd_data3a, liqhd_data3b) -> master_liqdata

master_liqdata %>%
  mutate(datetime = ymd_hms(datetime)) -> master_liqdata



master_liqdata <- read_csv("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files/master_liqhd_data.csv")

write_csv(master_liqdata, file = "master_liqhd_data.csv")

```
Testing out Matricies

```{r}
matrix(c(1,2,3,4,5,6), byrow = T, nrow = 3,ncol = 2)
```

CREATING TOTALS TIBBLE

```{r}

tail(master_liqdata)

# Simply filter by date to get data for each 4 day period

master_liqdata %>%
  filter(datetime <= ymd_hms("2024-11-04 23:59:00")) %>%
  select(-Experimental_side, -millis, -datetime) %>%
  summarize(across(everything(), sum)) -> totals_11_4

master_liqdata %>%
  filter(datetime >= ymd_hms("2024-11-05 00:00:00") & datetime <= ymd_hms("2024-11-08 23:59:00")) %>%
  select(-Experimental_side, -millis, -datetime) %>%
  summarize(across(everything(), sum)) -> totals_11_8

master_liqdata %>%
  filter(datetime >= ymd_hms("2024-11-09 00:00:00") & datetime <= ymd_hms("2024-11-12 23:59:00")) %>%
  select(-Experimental_side, -millis, -datetime) %>%
  summarize(across(everything(), sum)) -> totals_11_12

master_liqdata %>%
  filter(datetime >= ymd_hms("2024-11-13 00:00:00") & datetime <= ymd_hms("2024-11-15 23:59:00")) %>%
  select(-Experimental_side, -millis, -datetime) %>%
  summarize(across(everything(), sum)) -> totals_11_15

master_liqdata %>%
  filter(datetime >= ymd_hms("2024-11-16 00:00:00") & datetime <= ymd_hms("2024-11-18 23:59:00")) %>%
  select(-Experimental_side, -millis, -datetime) %>%
  summarize(across(everything(), sum)) -> totals_11_18




###############################################################################################################################


### (OLD CODE)
liqhd_data1a %>%
  select(-Experimental_side, -millis, -datetime) %>%
  summarize(across(everything(), sum)) -> totals_1a

liqhd_data1b %>%
  select(-Experimental_side, -millis, -datetime) %>%
  summarize(across(everything(), sum)) -> totals_1b

liqhd_data1c %>%
  select(-Experimental_side, -millis, -datetime) %>%
  summarize(across(everything(), sum)) -> totals_1c

liqhd_data2a %>%
  select(-Experimental_side, -millis, -datetime) %>%
  summarize(across(everything(), sum, na.rm = T)) -> totals_2a


totals_1a + totals_1b + totals_1c -> grand_totals_1


```

(UNNECESSARY)
MANIPULATING TOTALS TIBBLE

```{r}

### Splitting totals between left and right sides

grand_total %>%
  select(matches("\\d*[13579]\\b")) -> totals_LEFT 

#write.csv(totals_LEFT, file = "totals_LEFT.csv")

grand_total %>%
  select(matches("\\d*[24680]\\b")) -> totals_RIGHT


# Making a loop to turn 1 x 36 tibble into 2 x 18

result = matrix(NA, nrow = 18, ncol = 2)
i <- 1
while(i <= 36) {
  new_row <- grand_total %>% 
    select(i, i + 1) %>% 
    unlist() %>% 
    unname()
  result[i %/% 2, ] <- new_row
  i <- i + 2
}

matrix(as.numeric(grand_total)[ 1:36], ncol = 2, byrow = T)


#write.csv(result, file = 'result.csv')
```





Final Step to make updated totals tibble

```{r}


pivot_longer(totals_11_4, cols = everything(), names_to = "names", values_to = "values") %>%
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% 
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(Date = ymd("2024-11-04")) -> totals_11_4 

pivot_longer(totals_11_8, cols = everything(), names_to = "names", values_to = "values") %>%
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% 
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(Date = ymd("2024-11-08")) -> totals_11_8

pivot_longer(totals_11_12, cols = everything(), names_to = "names", values_to = "values") %>%
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% 
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(Date = ymd("2024-11-12")) -> totals_11_12

pivot_longer(totals_11_15, cols = everything(), names_to = "names", values_to = "values") %>%
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% 
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(Date = ymd("2024-11-15")) -> totals_11_15

pivot_longer(totals_11_18, cols = everything(), names_to = "names", values_to = "values") %>%
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% 
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(Date = ymd("2024-11-18")) -> totals_11_18 



### OLD CODE
### Organized Totals Tibble

pivot_longer(grand_totals_1, cols = everything(), names_to = "names", values_to = "values") %>%
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% # parse_number extracts any numbers found in column name
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(Date = ymd("2024-11-08")) -> updated_totals_1 # gsub removes any numbers in the column name

pivot_longer(totals_2a, cols = everything(), names_to = "names", values_to = "values") %>%
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% # parse_number extracts any numbers found in column name
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(Date = ymd("2024-11-12"))-> updated_totals_2

```




Creating bottle weight difference column using lag() function

```{r}

bottle_weights %>% 
  mutate(Date = mdy(Date)) %>%
  group_by(`Cage #`) %>% 
  mutate(weight_diff_left = -1 * (`Left Bottle` - lag(`Left Bottle`, 1))) %>% 
  mutate(weight_diff_right = -1 *(`Right Bottle` - lag(`Right Bottle`, 1))) %>% 
  rename(mouse_id = `Cage #`) -> bottle_weights

# I am filtering for only dates we are interested in
bottle_weights %>%
  filter(Date >= "2024-11-08") %>%
  filter(Action != "Refill") -> bottle_weights # Remove refilled bottle data because we don't need it


# Created extra column to differentiate the experimental side

bottle_weights %>%
  mutate(exp_side = "L") %>%
  select(Date, `mouse_id`, exp_side, `Left Bottle`, weight_diff_left) %>% 
  rename(bottle_weights = `Left Bottle`, weight_diff = weight_diff_left) -> bottle_weights_left
bottle_weights %>%
  mutate(exp_side = "R") %>%
  select(Date, `mouse_id`, exp_side,  `Right Bottle`, weight_diff_right) %>% 
  rename(bottle_weights = `Right Bottle`, weight_diff = weight_diff_right) -> bottle_weights_right

### This makes it so that bottle weight and weight diff have their own column. 

bind_rows(bottle_weights_left, bottle_weights_right) %>% group_by(mouse_id) -> bottle_weights
```





Create a column that specifies experimental side

```{r}

# Adding column that specifies Left and Right side

totals_11_8 %>% 
  mutate(exp_side = ifelse(parse_number(names) %% 2 == 0, "R", "L")) %>%
  select(Date, mouse_id, exp_side, measure_type, values) -> totals_11_8

totals_11_12 %>% 
  mutate(exp_side = ifelse(parse_number(names) %% 2 == 0, "R", "L")) %>%
  select(Date, mouse_id, exp_side, measure_type, values) -> totals_11_12

totals_11_15 %>% 
  mutate(exp_side = ifelse(parse_number(names) %% 2 == 0, "R", "L")) %>%
  select(Date, mouse_id, exp_side, measure_type, values) -> totals_11_15

totals_11_18 %>% 
  mutate(exp_side = ifelse(parse_number(names) %% 2 == 0, "R", "L")) %>%
  select(Date, mouse_id, exp_side, measure_type, values) -> totals_11_18





# OLD CODE
### After turning OG data set into a tidy format, split totals between left and right
updated_totals %>% 
  mutate(exp_side = "R") %>%
  filter(parse_number(names) %% 2 == 0) %>% 
  select(mouse_id, exp_side, measure_type, values) -> upd_totals_RIGHT

updated_totals %>% 
  mutate(exp_side = "L") %>%
  filter(parse_number(names) %% 2 != 0) %>% 
  select(mouse_id, exp_side, measure_type, values)-> upd_totals_LEFT

### Join Left and Right sides together
bind_rows(upd_totals_LEFT, upd_totals_RIGHT) %>%
  group_by(mouse_id) -> upd_totals_mod

### FASTER METHOD
# Instead of splitting, I can add exp side directly based on the number inside names

updated_totals_1 %>% 
  mutate(exp_side = ifelse(parse_number(names) %% 2 == 0, "R", "L")) %>%
  select(Date, mouse_id, exp_side, measure_type, values) -> updated_totals_1

updated_totals_2 %>% 
  mutate(exp_side = ifelse(parse_number(names) %% 2 == 0, "R", "L")) %>%
  select(Date, mouse_id, exp_side, measure_type, values) -> updated_totals_2




```


Slicing the data frame for the part we are interested in (LickNumber) 

```{r}
# Filtering each totals dataset for just lick number values
totals_11_8 %>%
  filter(measure_type == "LickNumber") -> totals_11_8_licknum

totals_11_12 %>%
  filter(measure_type == "LickNumber") -> totals_11_12_licknum

totals_11_15 %>%
  filter(measure_type == "LickNumber") -> totals_11_15_licknum

totals_11_18 %>%
  filter(measure_type == "LickNumber") -> totals_11_18_licknum

# Putting ONLY Lick Numbers in a single tibble
rbind(totals_11_8_licknum, totals_11_12_licknum, totals_11_15_licknum, totals_11_18_licknum) -> licknums

# joining bottle weights and lick nums
left_join(licknums, bottle_weights) -> totlicknum_and_bottle_weight


# OLD CODE
### Slicing data frames for part we are interested in
updated_totals_1 %>%
  filter(measure_type == "LickNumber") -> updated_totals_LICKNUM

# Totals from the next 4 day period
updated_totals_2 %>%
  filter(measure_type == "LickNumber") -> updated_totals2_LICKNUM

# Combining with prev. period
bind_rows(updated_totals_LICKNUM, updated_totals2_LICKNUM) -> LICKNUM


# Filtering bottle weights for the dates we are interested in
bottle_weights %>%
  filter(Date >= "2024-11-08" & Date <= "2024-11-12")-> bottle_weights11.08

#left_join(upd_totals_mod_LICKNUM, bottle_weights11.08) -> totlicknum_and_bottle_weight


# New code
# TODO: Problems with this code 

left_join(LICKNUM, bottle_weights11.08) -> totlicknum_and_bottle_weight
```


Plotting scatter plot between Total Lick Number and Bottle Weight Change


```{r}
### Plotting Data

lick_num_bottle_weight <- totlicknum_and_bottle_weight %>% 
  ggplot(aes(y = values,
             x = weight_diff, color = exp_side)) +
  geom_jitter() +
  #geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Lick Number Vs. Bottle Weight Difference",
       x = "Bottle Weight Difference (g)",
       y = "Number of Licks") +
  theme_bw()


lick_num_bottle_weight
ggsave("licknum_bottleweight_dot.png", plot = lick_num_bottle_weight)

```
```{r}
lm <- lm(values ~ weight_diff, data = totlicknum_and_bottle_weight)
summary(lm)
get_regression_table(lm)
```



Starting Creation of the Lick Duration vs. Weight Difference Plot

Filtering totals data for what we are interested in (LickDuration)
```{r}
# Filtering for Lick Duration only
totals_11_8 %>%
  filter(measure_type == "LickDuration") -> totals_11_8_lickdur

totals_11_12 %>%
  filter(measure_type == "LickDuration") -> totals_11_12_lickdur

totals_11_15 %>%
  filter(measure_type == "LickDuration") -> totals_11_15_lickdur

totals_11_18 %>%
  filter(measure_type == "LickDuration") -> totals_11_18_lickdur

# Combining the lickdur tibbles into one 
rbind(totals_11_8_lickdur, totals_11_12_lickdur, totals_11_15_lickdur, totals_11_18_lickdur) -> lickduration

# join with bottle weights data
left_join(lickduration, bottle_weights) -> totlickdur_and_bottle_weight

```
Plotting the relationship between lick duration and bottle weight difference

```{r}
lick_dur_bottle_weight <- totlickdur_and_bottle_weight %>% 
  ggplot(aes(y = values,
             x = weight_diff, color = exp_side)) +
  geom_jitter() +
  #geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Lick Duration Vs. Bottle Weight Difference",
       x = "Bottle Weight Difference (g)",
       y = "Lick Duration (ms)") +
  theme_bw()


lick_dur_bottle_weight
ggsave("lickdur_bottleweight_dot.png", plot = lick_dur_bottle_weight)
```





Scatter plot for Lick Number and Lick Duration

```{r}

head(lickduration)
licknums %>%
  rename(licknums = values) %>%
  select(-measure_type) -> licknums

lickduration %>%
  rename(lickduration = values)%>%
  select(-measure_type) -> lickduration

left_join(licknums, lickduration) -> licknums_and_dur
```
Scatter plot

```{r}
licknum_dur <- licknums_and_dur %>% 
  ggplot(aes(y = licknums,
             x = lickduration)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Lick Number Vs. Lick Duration",
       x = "Total Lick Duration (sec)",
       y = "Total Lick Number") +
  theme_bw()


licknum_dur
ggsave("licknum_dur_mod.png", plot = licknum_dur)

```
Extracting ONLY BOUT NUMBER

```{r}
totals_11_8 %>%
  filter(measure_type == "BoutNumber") -> totals_11_8_boutnum

totals_11_12 %>%
  filter(measure_type == "BoutNumber") -> totals_11_12_boutnum

totals_11_15 %>%
  filter(measure_type == "BoutNumber") -> totals_11_15_boutnum

totals_11_18 %>%
  filter(measure_type == "BoutNumber") -> totals_11_18_boutnum

rbind(totals_11_8_boutnum, totals_11_12_boutnum, totals_11_15_boutnum, totals_11_18_boutnum) -> boutnum


```

Extracting ONLY BOUT DURATION

```{r}
totals_11_8 %>%
  filter(measure_type == "BoutDuration") -> totals_11_8_boutdur

totals_11_12 %>%
  filter(measure_type == "BoutDuration") -> totals_11_12_boutdur

totals_11_15 %>%
  filter(measure_type == "BoutDuration") -> totals_11_15_boutdur

totals_11_18 %>%
  filter(measure_type == "BoutDuration") -> totals_11_18_boutdur

rbind(totals_11_8_boutdur, totals_11_12_boutdur, totals_11_15_boutdur, totals_11_18_boutdur) -> boutdur

```

Combining bout num and bout dur

```{r}
boutnum %>%
  rename(boutnums = values) %>%
  select(-measure_type) -> boutnum

boutdur %>%
  rename(boutduration = values)%>%
  select(-measure_type) -> boutdur

left_join(boutnum, boutdur) -> boutnum_boutdur
```
Plotting Bout Number vs Bout Duration

```{r}
boutnum_boutdur_plot <- boutnum_boutdur %>% 
  ggplot(aes(y = boutnums,
             x = boutduration)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Bout Number Vs. Bout Duration",
       x = "Total Bout Duration (sec)",
       y = "Total Bout Number") +
  theme_bw()


boutnum_boutdur_plot
ggsave("boutnum_boutdur_plot.png", plot = boutnum_boutdur_plot)
```





