---
title: "AQ_NEW_liqhd_preference"
output: html_document
date: "2025-02-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(lubridate)
library(dplyr)
library(dbplot)
library(stringr)
library(tidyr)
library(GGally)
library(skimr)
library(moderndive)
library(readxl)

# Lab PC
knitr::opts_knit$set(root.dir = normalizePath("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files"))

# For PC
knitr::opts_knit$set(root.dir = normalizePath("C:/Users/antho/OneDrive - University of Massachusetts/Richardson Lab/Anthony R Files/Data Files"))

# For Laptop
knitr::opts_knit$set(root.dir = normalizePath("C:/Users/Thinkpad-X1/OneDrive - University of Massachusetts/Richardson Lab/Anthony R Files/Data Files"))
```

Importing new data

```{r}
library(readr)
data_1.28 <- read_csv("25012807.csv", col_select = -...220) 
# For some reason data file always comes with random "...220" column full of NAs
```


Pivoting Data into longer format 

```{r}
head(data_1.28)


# Turning datetime column into actual date-type object
data_1.28 %>% 
  mutate(datetime = ymd_hms(datetime)) -> data_1.28

# Pivoting the entire data set into longer format
data_1.28 %>% 
  select(-Experimental_side, -millis) %>% # Remove the cols we don't want
  pivot_longer(!datetime, values_to = "values", names_to = "names") %>% 
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% 
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(exp_side = ifelse(parse_number(names) %% 2 == 0, "R", "L")) %>% 
  select(datetime, mouse_id, exp_side, measure_type, values) -> longer_data


```

Creating Hour Column

```{r}
longer_data %>%
  mutate(hour = floor_date(datetime, "hour")) %>% 
  select(datetime, hour, everything()) -> longer_data

```


Filtering longer data for only mice we want (1-6) AND classifying each mouse empty vs experimental

```{r}
longer_data %>%
  filter(mouse_id <= 6) %>%
  mutate(treatment = ifelse(mouse_id == 6, "Empty", "Experimental")) %>% 
  select(datetime, hour, treatment, everything()) -> longer_data_filt
```

Turning lick numbers into proportions
  a) Finding Total number of licks (regardless of exp side)
```{r}
# finding total licks per hour for each mice (left + right lick nums)
longer_data_filt %>% 
  filter(measure_type == "LickNumber") %>%
  group_by(hour, mouse_id, measure_type, treatment) %>%
  summarize(total_value = sum(values)) -> licks_total 


#right_join(licks_longer_filt, licks_total) -> licks_total # Adding totals column to longer dataset 

#licks_total %>%
  #mutate(lick_prop = round((values/total_value), 4)) -> licks_prop # Dividing values col by total values col to get proportion

```
  b) Finding total number of LEFT (experimental side) licks 

```{r}
longer_data_filt %>%
  filter(measure_type == "LickNumber") %>%
  filter(exp_side == "L") %>% 
  group_by(hour, mouse_id) %>% 
  summarize(left_licks = sum(values)) -> left_licks_total

```
  c) Combining the previous 2 tibbles to get left lick proportion
  
```{r}
left_join(left_licks_total, licks_total) %>% 
  mutate(lick_prop = round((left_licks/total_value), 4)) -> licks_prop_hour
```


Plotting Left Preference

```{r}
pref_chart <- ggplot(licks_prop_hour, aes(x = hour, y = lick_prop)) +
  geom_line(aes(color = treatment), size = 1.2, alpha = 0.8) +
  facet_wrap(~ mouse_id, scales = "free_y") +
  geom_line(aes(y = 0.5), color = "black") +
  labs(title = "Left Side Preference (Experimental vs. Empty Cages)",
       subtitle = "Separated by Mouse ID",
       color = "Treatment Group",
       x = "Time (Hours)",
       y = "Preference %") +
    scale_x_datetime(date_breaks = "2 hours", date_labels = "%H:%M") +  # Label every 2 hours
  #theme_minimal(base_size = 14) + 
  theme(panel.grid.major = element_line(color = "grey80", linetype = "dashed"),
        panel.grid.minor = element_blank(),
        legend.position = "top",
        strip.background = element_rect(fill = "lightgrey", color = "black"),
        axis.text.x = element_text(angle = 45, hjust = 1))

  #facet_wrap(licks_prop_hour$mouse_id)

pref_chart
ggsave("NEW_preference_chart.png", plot = pref_chart)
# Problem with the chart is that some points in data have 0s for value, thus making a proportion greater than 1
```

