---
title: "AQ_Preference_Chart"
output: html_document
date: "2024-11-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(lubridate)
library(dplyr)
library(dbplot)
library(stringr)
library(tidyr)
library(GGally)
library(skimr)
library(moderndive)
library(readxl)
 
#setwd("C:/Users/lab/OneDrive - University of Massachusetts (2)/Anthony R Files/Data Files")
```

Importing Data

```{r}
library(readr)
master_liqhd_data <- read_csv("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files/master_liqhd_data.csv")
tail(master_liqhd_data)

master_liqhd_data <- read_csv("master_liqhd_data.csv")


#liqhd_data3c <- read_csv("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files/24112001.CSV")

#liqhd_data3d <- read_csv("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files/24112200.CSV",  col_select = -...220)

#liqhd_data4a <- read_csv("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files/24112500.CSV",  col_select = -...220)

#liqhd_data4b <- read_csv("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files/24112700.CSV",  col_select = -...220)

liqhd_data5a <- read_csv("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files/24110800.CSV",  col_select = -...220)

#liqhd_data3c %>%
  mutate(datetime = ymd_hms(datetime)) -> liqhd_data3c

#liqhd_data3d %>%
  mutate(datetime = ymd_hms(datetime)) -> liqhd_data3d

#liqhd_data4a %>%
  mutate(datetime = ymd_hms(datetime)) -> liqhd_data4a

#liqhd_data4b %>%
  mutate(datetime = ymd_hms(datetime)) -> liqhd_data4b

# Added new data to master tibble
rbind(master_liqhd_data, liqhd_data5a) -> master_liqhd_data




### COMBINES ALL NEW DATA WITH OLD
write_csv(master_liqhd_data, file = "master_liqhd_data.csv")

```


Manipulating master data tibble
 
```{r}
master_liqhd_data %>%
  select(-Experimental_side, -millis) -> master_liqhd_data_values

master_liqhd_data_values %>% 
  pivot_longer(!datetime, values_to = "values", names_to = "names") -> master_liqhd_longer

```
 
Modifying the long data by adding columns for name, exp side, and mouse id. Basically everything we need to be able to differentiate every row.

```{r}
master_liqhd_longer %>%
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% 
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(exp_side = ifelse(parse_number(names) %% 2 == 0, "R", "L")) %>%
  select(datetime, mouse_id, exp_side, measure_type, values) -> master_liqhd_longer


# Parsing out an hour, day, month column. I am going to be using this to condense the data down for every hour rather than every minute

#master_liqhd_longer %>%
  #mutate(hour = hour(datetime)) %>%
  #mutate(day = day(datetime)) %>%
  #mutate(month = month(datetime)) %>%
  #mutate(year = year(datetime)) -> master_liqhd_longer

# Simplified code
master_liqhd_longer %>%
  mutate(hour = floor_date(datetime, "hour")) -> master_liqhd_longer
```

Condensing master_liqhd_longer down to one data observation per hour

```{r}

# This tibble has total licks SEPARATED between left and right sides
master_liqhd_longer %>%
  filter(measure_type == "LickNumber") %>%
  filter(mouse_id != 18) %>%
  group_by(hour, exp_side) %>%
  summarize(total_value = sum(values)) -> liqhd_licks_hour

# This tibble is half the size of the one right above
# This contains an ONLY TOTAL VALUES column that adds up values regardless of exp side
master_liqhd_longer %>%
  filter(measure_type == "LickNumber") %>%
  filter(mouse_id != 18) %>%
  group_by(hour) %>%
  summarize(total_value = sum(values)) -> liqhd_licks_hour_total
```

Converting my hourly totals tibble into proportions instead of numbers

```{r}
left_join(liqhd_licks_hour, liqhd_licks_hour_total, by = c("hour")) %>%
  rename(value = total_value.x) %>%
  rename(total_value = total_value.y) %>%
  mutate(value = value/total_value) -> licks_by_hour_props
```

(UNNECESSARY)
Starting the preference chart
Separating Left and Right

```{r}
#licks_by_hour_props %>%
  #filter(exp_side == "L") -> licks_hour_props_LEFT

#licks_by_hour_props %>%
  #filter(exp_side == "R") -> licks_hour_props_RIGHT


```




```{r}
# Left Preference ONLY

#preference_chart <- licks_hour_props_LEFT %>%
  #ggplot() +
  #geom_point(mapping = aes(x = hour, y = value))

#preference_chart

# Right Preference ONLY

#preference_chart <- licks_hour_props_RIGHT %>%
 #ggplot() +
  #geom_point(mapping = aes(x = hour, y = value)) 

#preference_chart

# Preference for BOTH
preference_chart <- licks_by_hour_props %>%
  ggplot() +
  geom_line(mapping = aes(x = hour, y = value, color = exp_side), size = 0.75) +
  labs(title = "Preference % Left vs. Right Side",
       x = "Date",
       y = "Side Preference (%)",
       color = "Experimental Side") +
  theme_bw() +
  scale_x_datetime(date_breaks = "1440 min", date_labels = "%m/%d") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_line(aes(y = 0.5))

preference_chart

#ggsave("preference_chart.png", plot = preference_chart)
#which(licks_by_hour_props$value <= 0)
#licks_by_hour_props[3053,]
```


