---
title: "AQ_Preference_Chart_Updated"
output: html_document
date: "2024-12-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(lubridate)
library(dplyr)
library(dbplot)
library(stringr)
library(tidyr)
library(GGally)
library(skimr)
library(moderndive)
library(readxl)
```


(OLD) Importing Data

```{r}
# Set working directory as the Data Files folder
library(readr)
#filt_master_liq_long <- read_csv("filt_master_liqhd_longer.csv")

liqhd_data_11_12 <- read_csv("24111201.CSV")
liqhd_data_11_18 <- read_csv("24111800.CSV")
liqhd_data_11_20 <- read_csv("24112000.CSV")
liqhd_data_11_201 <- read_csv("24112001.CSV")
liqhd_data_11_22 <- read_csv("24112200.CSV",  col_select = -...220)
liqhd_data_11_25 <- read_csv("24112500.CSV",  col_select = -...220)
liqhd_data_11_27 <- read_csv("24112700_edited.CSV")

rbind(liqhd_data_11_12, liqhd_data_11_18, liqhd_data_11_20, liqhd_data_11_201, liqhd_data_11_22, liqhd_data_11_25, liqhd_data_11_27) -> master_liqhd_data

write_csv(master_liqhd_data, file = "master_liqhd_data.csv")
```

Importing Master Data Set
```{r}
library(readr)
master_liqhd_data <- read_csv("master_liqhd_data.csv")

```

Making Master Dataset Longer

```{r}
master_liqhd_data %>% 
  select(-Experimental_side, -millis) %>% 
  pivot_longer(!datetime, values_to = "values", names_to = "names") %>% 
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% 
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(exp_side = ifelse(parse_number(names) %% 2 == 0, "R", "L")) %>% 
  select(datetime, mouse_id, exp_side, measure_type, values) -> master_liqhd_longer

```

(OLD) Filtering the imported data to remove empty cages

```{r}
liqhd_data_11_12 %>% 
  select(-matches("27|28|35|36")) %>% # remove cage 14, 18
  select(-Experimental_side, -millis) %>% 
  pivot_longer(!datetime, values_to = "values", names_to = "names") %>%
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% 
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(exp_side = ifelse(parse_number(names) %% 2 == 0, "R", "L")) %>%
  select(datetime, mouse_id, exp_side, measure_type, values) -> liqhd_data_11_12
  

liqhd_data_11_18 %>% 
  select(-matches("35|36")) %>% # remove cage 18
  select(-Experimental_side, -millis) %>% 
  pivot_longer(!datetime, values_to = "values", names_to = "names") %>%
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% 
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(exp_side = ifelse(parse_number(names) %% 2 == 0, "R", "L")) %>%
  select(datetime, mouse_id, exp_side, measure_type, values) -> liqhd_data_11_18

liqhd_data_11_20 %>% 
  select(-matches("27|28|35|36")) %>% # remove cage 14, 18
  select(-Experimental_side, -millis) %>% 
  pivot_longer(!datetime, values_to = "values", names_to = "names") %>%
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% 
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(exp_side = ifelse(parse_number(names) %% 2 == 0, "R", "L")) %>%
  select(datetime, mouse_id, exp_side, measure_type, values) -> liqhd_data_11_20

liqhd_data_11_201 %>% 
  select(-matches("27|28|35|36")) %>% # remove cage 14, 18
  select(-Experimental_side, -millis) %>% 
  pivot_longer(!datetime, values_to = "values", names_to = "names") %>%
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% 
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(exp_side = ifelse(parse_number(names) %% 2 == 0, "R", "L")) %>%
  select(datetime, mouse_id, exp_side, measure_type, values) -> liqhd_data_11_201

liqhd_data_11_22 %>% 
  select(-matches("27|28|35|36")) %>% # Remove cage 14, 18
  select(-Experimental_side, -millis) %>% 
  pivot_longer(!datetime, values_to = "values", names_to = "names") %>%
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% 
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(exp_side = ifelse(parse_number(names) %% 2 == 0, "R", "L")) %>%
  select(datetime, mouse_id, exp_side, measure_type, values) -> liqhd_data_11_22

liqhd_data_11_25 %>% 
  select(-matches("15|16|27|28|35|36")) %>% # remove cage 8, 14, 18
  select(-Experimental_side, -millis) %>% 
  pivot_longer(!datetime, values_to = "values", names_to = "names") %>%
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% 
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(exp_side = ifelse(parse_number(names) %% 2 == 0, "R", "L")) %>%
  select(datetime, mouse_id, exp_side, measure_type, values) -> liqhd_data_11_25

liqhd_data_11_27 %>% 
  select(-matches("7|8|15|16|27|28|35|36")) %>% 
  select(-Experimental_side, -millis) %>% 
  pivot_longer(!datetime, values_to = "values", names_to = "names") %>%
  mutate(mouse_id = (parse_number(names) + 1) %/% 2 ) %>% 
  mutate(measure_type = gsub("[[:digit:]]", "", names)) %>%
  mutate(exp_side = ifelse(parse_number(names) %% 2 == 0, "R", "L")) %>%
  select(datetime, mouse_id, exp_side, measure_type, values) -> liqhd_data_11_27

rbind(liqhd_data_11_12, liqhd_data_11_18, liqhd_data_11_20,liqhd_data_11_201, liqhd_data_11_22, liqhd_data_11_25, liqhd_data_11_27) -> filt_master_liq_longer

write_csv(filt_master_liq_longer, file = "filt_master_liqhd_longer.csv")

  
```

Filtering Longer Data to remove empty cages AND making hour column

```{r}
master_liqhd_longer %>%
  filter(!(mouse_id %in% c(4,8,14,18))) %>%
  mutate(datetime = ymd_hms(datetime)) %>%
  mutate(hour = floor_date(datetime, "hour"))-> filt_liqhd_longer
```

(OLD) Creating Hour Column

```{r}
filt_master_liq_long %>%
  mutate(datetime = ymd_hms(datetime)) %>%
  mutate(hour = floor_date(datetime, "hour")) -> filt_master_liq_long

head(filt_master_liq_long)
```

Creating a Treatment Column

```{r}
filt_liqhd_longer %>%
  mutate(treatment = if_else(mouse_id %in% c(1,3,6,7,11,14,15,16), "Water", "Alchohol")) %>%
  select(datetime, hour, treatment, everything()) -> filt_liqhd_longer
```

Filtering only for Lick Number & Turning values into proportions

```{r}
filt_liqhd_longer %>%
  filter(measure_type == "LickNumber") %>%
  filter(exp_side == "R") %>%
  group_by(hour, exp_side, treatment) %>%
  summarize(total_value = sum(values)) %>%
  rename(value = total_value) -> liqhd_licks_hour

filt_liqhd_longer %>% 
  filter(measure_type == "LickNumber") %>%
  group_by(hour) %>%
  summarize(total_value = sum(values)) -> liqhd_licks_hour_total

left_join(liqhd_licks_hour, liqhd_licks_hour_total) %>%
  mutate(value = value/total_value)->  licks_by_hour_props
 
```

(OLD) Preference Chart

```{r}
upd_preference_chart <- licks_by_hour_props %>%
  ggplot() +
  geom_line(mapping = aes(x = hour, y = value, color = exp_side), size = 0.75) +
  labs(title = "Preference % Left vs. Right Side",
       x = "Date",
       y = "Side Preference (%)",
       color = "Experimental Side") +
  theme_bw() 
 
upd_preference_chart

ggsave("upd_preference_chart.png", plot = upd_preference_chart)
```
(OLD) Another chart
```{r}
preference_chart2 <- 
  licks_by_hour_props %>%
  ggplot() +
  geom_line(mapping = aes(x = hour, y = value, color = mouse_id), size = 0.75) +
  labs(title = "Preference % Left vs. Right Side",
       x = "Date",
       y = "Side Preference (%)",
       color = "Mouse ID") +
  theme_bw() 

preference_chart2

ggsave("preference_chart2.png", plot = preference_chart2)
 
```

(USELESS) Separating Water and Alcohol treatment groups

```{r}
licks_by_hour_props %>%
  filter(treatment == "Water") -> licks_by_hour_props_WATER

licks_by_hour_props %>% 
  filter(treatment == "Alcohol") -> licks_by_hour_props_ALC


```

Right Side Preference chart separated by treatment group

```{r}
preference_chart <- ggplot(licks_by_hour_props, aes(x = hour, y = value)) +
  geom_line(size = 0.7, aes(color = treatment)) +
   scale_x_datetime(date_breaks = "1440 min", date_labels = "%m/%d") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_line(aes(y = 0.5), color = "black") +
  labs(title = "Right Side Preference (Alcohol vs. Water Group)",
       x = "Time (Hours)",
       y = "Preference %",
       color = "Treatment Group") +
  geom_rect(aes(xmin = as.POSIXct("2024-11-15 06:00:00"),
                xmax = as.POSIXct("2024-11-18 06:00:00"),
                ymin = -Inf,
                ymax = Inf),
            alpha = 0.01) 


preference_chart 

ggsave("preference_chart_treatment.png", plot = preference_chart)

```

Preparing lick props for for facet wrap (Same code as above, but includes mouse_id in group_by)

```{r}
filt_liqhd_longer %>%
  filter(measure_type == "LickNumber") %>%
  filter(exp_side == "R") %>%
  group_by(hour, exp_side, treatment, mouse_id) %>%
  summarize(total_value = sum(values)) %>%
  rename(value = total_value) -> liqhd_licks_hour

filt_liqhd_longer %>% 
  filter(measure_type == "LickNumber") %>%
  group_by(hour, mouse_id) %>%
  summarize(total_value = sum(values)) -> liqhd_licks_hour_total

left_join(liqhd_licks_hour, liqhd_licks_hour_total) %>%
  mutate(value = value/total_value)->  licks_by_hour_props
```


Using facet_wrap to see individual mouse Right side Preference

```{r}
facet_wrap_pref_chart <- 
  ggplot(licks_by_hour_props,
         aes(x = hour,
             y = value)) +
  geom_line(aes(color = treatment)) +
  facet_wrap(licks_by_hour_props$mouse_id) +
  scale_x_datetime(date_breaks = "5760 min",
                   date_labels = "%m/%d") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  geom_line(aes(y = 0.5),
            color = "black") +
  labs(title = "Right Side Preference (Alcohol vs. Water Group)",
       subtitle = "Separated by Mouse ID",
       color = "Treatment Group",
       x = "Time (Hours)",
       y = "Preference %") +
   geom_rect(alpha = 0.01, aes(xmin = as.POSIXct("2024-11-15 12:00:00"),
                xmax = as.POSIXct("2024-11-18 12:00:00"),
                ymin = -Inf,
                ymax = Inf))

facet_wrap_pref_chart

ggsave("facet_wrap_pref_chart.png", plot = facet_wrap_pref_chart)

```


