---
title: "Cell_counts_Ranalysis_SOP"
author: "Anthony Qiu"
date: "2025-04-10"
output: html_document
---

This block is a package check. Think of these as "mods" to help you do your analysis
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) # Adds all data manipulation and visualization tools
library(ggpubr) # Uses ggplot to make publication-ready visualizations
#library(RColorBrewer) # Adds color palettes 
#library(Routliers) # Outlier detection package
```


Download data here. You need a different path depending on which device you are using. 
```{r}
# For my personal laptop
cell_counts <- read_csv("C:/Users/Thinkpad-X1/OneDrive - University of Massachusetts/Richardson Lab/Anthony R Files/Data Files/CREST20X_data_6.csv")

# Lab PC
cell_counts <- read_csv("C:/Users/lab/OneDrive - University of Massachusetts (2)/Richardson Lab/Anthony R Files/Data Files/CREST20X_data_6.csv")

# My personal PC
cell_counts <- read_csv("C:/Users/antho/OneDrive - University of Massachusetts/Richardson Lab/Anthony R Files/Data Files/CREST20X_data_6.csv")


```


(STEP 1a) Averaging cell density between left and right hemispheres
```{r}
# Avg between hemispheres for DAPI, CC1, ASPA density
cell_counts %>%
  group_by(Season, `Bregma(mm)`, ROI, MouseID) %>%
  summarize(mean(TotalDAPIDensitymm2), mean(TotalCC1Densitymm2), mean(TotalASPADensitymm2)) -> avg_density

# Renaming the columns for cleanliness
avg_density %>%
  rename(DAPIdensity = `mean(TotalDAPIDensitymm2)`) %>%
  rename(CC1density = `mean(TotalCC1Densitymm2)`) %>%
  rename(ASPAdensity = `mean(TotalASPADensitymm2)`) -> avg_density

```


(STEP 1b) Manipulate tibble to longer format for easier visualization
```{r}
long_data <- avg_density %>%
  pivot_longer(cols = c(DAPIdensity, CC1density, ASPAdensity), # Adjust these if different
               names_to = "Cell_Type",
               values_to = "Densities")

```



(STEP 2) Creating Median, MAD, Upper & Lower Bound, and Outlier columns using dplyr to manipulate original data frame. 

What this does is that it first groups by Season, Bregma, ROI, and Cell Type and finds Median and MAD for each group. 
 - What this would mean for our graph is that this code block computes Median and MAD for every column. 
 
Then, we combine the resulting dataframe with our original "long_data" so that we can have the Median and MAD columns in the same data frame as the rest of our data; This just makes calculating outliers easier

Then because we want a poorly conservative interval, this means that we would be LESS strict on what counts as an outlier. 
```{r}
long_data %>%
  group_by(Season, `Bregma(mm)`, ROI, Cell_Type) %>%
  summarize(Median = median(Densities), MAD = mad(Densities)) %>%
  left_join(.,long_data) %>%
  mutate(Upper_Bound = Median + 2 * MAD) %>%
  mutate(Lower_Bound = Median - 2 * MAD) %>%
  mutate(Outlier = ifelse(Densities < Lower_Bound | Densities > Upper_Bound, 1, 0)) -> all_outlier_info
```



(STEP 4) Creating an interaction group so that I can have a column per bregma & treatment group combination
```{r}
all_outlier_info %>%
   mutate(bregma_group = interaction(`Bregma(mm)`, Season, sep = "mm | ")) -> all_outlier_info
```


(STEP 5a) Plot cell densities for Corpus Callosum Region of Interest
(DAPI)
```{r}
dapi_cc <- ggplot(data = all_outlier_info %>%
                    filter(Cell_Type == "DAPIdensity") %>%
                    filter(ROI == "CC"),
                  aes(x = bregma_group, y = Densities)) +
  
  geom_col(position = "dodge", aes(y = Median), fill = NA, color = "black", linewidth = 1, width = 0.75) +
  geom_errorbar(aes(ymin = Lower_Bound, ymax = Upper_Bound), linewidth = 1, width = 0.5) +
  geom_jitter(data = all_outlier_info %>%
                filter(Cell_Type == "DAPIdensity") %>%
                filter(ROI == "CC") %>%
                filter(Outlier == 1),
              width = 0.2,
              shape = 18,
              size = 2,
              color = "red") +
  geom_jitter(data = all_outlier_info %>%
                filter(Cell_Type == "DAPIdensity") %>%
                filter(ROI == "CC") %>%
                filter(Outlier == 0),
              width = 0.2,
              size = 2) +
  
  labs(title = "Total DAPI Density by Group in Corpus Callosum",
         x = "Bregma & Season",
         y = "Densities (mm^2)") +
  
  theme_pubr() +
  theme(plot.title = element_text(face = "bold", size = 14),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text.x = element_text(face = "bold", angle = 45, hjust = 1),
        axis.text.y = element_text(face = "bold"),
        axis.line = element_line(linewidth = 1, color = "black")) +
  coord_cartesian(ylim = c(3500, NA ))

dapi_cc
```


(STEP 5b) Corpus Callosum
(CC1)
```{r}
cc1_cc <- ggplot(data = all_outlier_info %>%
                    filter(Cell_Type == "CC1density") %>%
                    filter(ROI == "CC"),
                  aes(x = bregma_group, y = Densities)) +
  
  geom_col(position = "dodge", aes(y = Median), fill = NA, color = "black", linewidth = 1, width = 0.75) +
  geom_errorbar(aes(ymin = Lower_Bound, ymax = Upper_Bound), linewidth = 1, width = 0.5) +
  geom_jitter(data = all_outlier_info %>%
                filter(Cell_Type == "CC1density") %>%
                filter(ROI == "CC") %>%
                filter(Outlier == 1),
              width = 0.2,
              shape = 18,
              size = 2,
              color = "red") +
  geom_jitter(data = all_outlier_info %>%
                filter(Cell_Type == "CC1density") %>%
                filter(ROI == "CC") %>%
                filter(Outlier == 0),
              width = 0.2,
              size = 2) +
  
  labs(title = "Total CC1 Density by Group in Corpus Callosum",
         x = "Bregma & Season",
         y = "Densities (mm^2)") +
  
  theme_pubr() +
  theme(plot.title = element_text(face = "bold", size = 14),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text.x = element_text(face = "bold", angle = 45, hjust = 1),
        axis.text.y = element_text(face = "bold"),
        axis.line = element_line(linewidth = 1, color = "black"))

cc1_cc

```


(STEP 5c) Corpus Callosum
(ASPA)
```{r}
aspa_cc <- ggplot(data = all_outlier_info %>%
                    filter(Cell_Type == "ASPAdensity") %>%
                    filter(ROI == "CC"),
                  aes(x = bregma_group, y = Densities)) +
  
  geom_col(position = "dodge", aes(y = Median), fill = NA, color = "black", linewidth = 1, width = 0.75) +
  geom_errorbar(aes(ymin = Lower_Bound, ymax = Upper_Bound), linewidth = 1, width = 0.5) +
  geom_jitter(data = all_outlier_info %>%
                filter(Cell_Type == "ASPAdensity") %>%
                filter(ROI == "CC") %>%
                filter(Outlier == 1),
              width = 0.2,
              shape = 18,
              size = 2,
              color = "red") +
  geom_jitter(data = all_outlier_info %>%
                filter(Cell_Type == "ASPAdensity") %>%
                filter(ROI == "CC") %>%
                filter(Outlier == 0),
              width = 0.2,
              size = 2) +
  
  labs(title = "Total ASPA Density by Group in Corpus Callosum",
         x = "Bregma & Season",
         y = "Densities (mm^2)") +
  
  theme_pubr() +
  theme(plot.title = element_text(face = "bold", size = 14),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text.x = element_text(face = "bold", angle = 45, hjust = 1),
        axis.text.y = element_text(face = "bold"),
        axis.line = element_line(linewidth = 1, color = "black"))

aspa_cc

```


(STEP 6a) Now repeat all of STEP 5 but with Cingulate Gyrus
(DAPI)
```{r}
dapi_cg <- ggplot(data = all_outlier_info %>%
                    filter(Cell_Type == "DAPIdensity") %>%
                    filter(ROI == "Cg"),
                  aes(x = bregma_group, y = Densities)) +
  
  geom_col(position = "dodge", aes(y = Median), fill = NA, color = "black", linewidth = 1, width = 0.75) +
  geom_errorbar(aes(ymin = Lower_Bound, ymax = Upper_Bound), linewidth = 1, width = 0.5) +
  geom_jitter(data = all_outlier_info %>%
                filter(Cell_Type == "DAPIdensity") %>%
                filter(ROI == "Cg") %>%
                filter(Outlier == 1),
              width = 0.2,
              shape = 18,
              size = 2,
              color = "red") +
  geom_jitter(data = all_outlier_info %>%
                filter(Cell_Type == "DAPIdensity") %>%
                filter(ROI == "Cg") %>%
                filter(Outlier == 0),
              width = 0.2,
              size = 2) +
  
  labs(title = "Total DAPI Density by Group in Cingulate Gyrus",
         x = "Bregma & Season",
         y = "Densities (mm^2)") +
  
  theme_pubr() +
  theme(plot.title = element_text(face = "bold", size = 14),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text.x = element_text(face = "bold", angle = 45, hjust = 1),
        axis.text.y = element_text(face = "bold"),
        axis.line = element_line(linewidth = 1, color = "black")) +
  scale_y_continuous(breaks = seq(0, 6500, by = 250)) +
  coord_cartesian(ylim = c(3500, NA))
 
dapi_cg
```

(STEP 6b) Cingulate Gyrus
(CC1)
```{r}
cc1_cg <- ggplot(data = all_outlier_info %>%
                    filter(Cell_Type == "CC1density") %>%
                    filter(ROI == "Cg"),
                  aes(x = bregma_group, y = Densities)) +
  
  geom_col(position = "dodge", aes(y = Median), fill = NA, color = "black", linewidth = 1, width = 0.75) +
  geom_errorbar(aes(ymin = Lower_Bound, ymax = Upper_Bound), linewidth = 1, width = 0.5) +
  geom_jitter(data = all_outlier_info %>%
                filter(Cell_Type == "CC1density") %>%
                filter(ROI == "Cg") %>%
                filter(Outlier == 1),
              width = 0.2,
              shape = 18,
              size = 2,
              color = "red") +
  geom_jitter(data = all_outlier_info %>%
                filter(Cell_Type == "CC1density") %>%
                filter(ROI == "Cg") %>%
                filter(Outlier == 0),
              width = 0.2,
              size = 2) +
  
  labs(title = "Total CC1 Density by Group in Cingulate Gyrus",
         x = "Bregma & Season",
         y = "Densities (mm^2)") +
  
  theme_pubr() +
  theme(plot.title = element_text(face = "bold", size = 14),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text.x = element_text(face = "bold", angle = 45, hjust = 1),
        axis.text.y = element_text(face = "bold"),
        axis.line = element_line(linewidth = 1, color = "black"))

cc1_cg

```

(STEP 6c) Cingulate Gyrus
(ASPA)
```{r}
aspa_cg <- ggplot(data = all_outlier_info %>%
                    filter(Cell_Type == "ASPAdensity") %>%
                    filter(ROI == "Cg"),
                  aes(x = bregma_group, y = Densities)) +
  
  geom_col(position = "dodge", aes(y = Median), fill = NA, color = "black", linewidth = 1, width = 0.75) +
  geom_errorbar(aes(ymin = Lower_Bound, ymax = Upper_Bound), linewidth = 1, width = 0.5) +
  geom_jitter(data = all_outlier_info %>%
                filter(Cell_Type == "ASPAdensity") %>%
                filter(ROI == "Cg") %>%
                filter(Outlier == 1),
              width = 0.2,
              shape = 18,
              size = 2,
              color = "red") +
  geom_jitter(data = all_outlier_info %>%
                filter(Cell_Type == "ASPAdensity") %>%
                filter(ROI == "Cg") %>%
                filter(Outlier == 0),
              width = 0.2,
              size = 2) +
  
  labs(title = "Total ASPA Density by Group in Corpus Callosum",
         x = "Bregma & Season",
         y = "Densities (mm^2)") +
  
  theme_pubr() +
  theme(plot.title = element_text(face = "bold", size = 14),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text.x = element_text(face = "bold", angle = 45, hjust = 1),
        axis.text.y = element_text(face = "bold"),
        axis.line = element_line(linewidth = 1, color = "black"))

aspa_cg

```

