---
title: "STATS310_Project"
output: pdf_document
date: "2024-12-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(infer)
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)
library(forcats)
library(readr)
```


Importing data 

```{r}
library(readr)
philly_data <- read_csv("pa_philadelphia_2020_04_01.csv")

nj_statewide_data <-  read_csv("nj_statewide_2020_04_01.csv")

# Creating county data files so its easier to load for next time. 
write_csv(gloucester_county_data, "gloucester_county_data.csv")
write_csv(cumberland_county_data, "cumberland_county_data.csv")
write_csv(capemay_county_data, "capemay_county_data.csv")
write_csv(salem_county_data, "salem_county_data.csv")
write_csv(camden_data, "camden_city_data.csv")

gloucester_county_data <- read_csv("gloucester_county_data.csv")
cumberland_county_data <- read_csv("cumberland_county_data.csv")
capemay_county_data <- read_csv("capemay_county_data.csv")
salem_county_data <- read_csv("salem_county_data.csv")
camden_data <- read_csv("camden_city_data.csv")


```
Filtering NJ data by the counties we want.
Because the dataset doesn't have a county column, we are grouping 3 of the largest townships (in terms of pop) in each county to represent that county.

```{r}
camden_data <- nj_statewide_data %>%
  filter(raw_TOWNSHIP == "Camden City")

gloucester_county_data <- nj_statewide_data %>%
  filter(raw_TOWNSHIP == "Washington Twp" |raw_TOWNSHIP == "Monroe Twp"  | raw_TOWNSHIP == "Deptford Twp")

cumberland_county_data <- nj_statewide_data %>%
  filter(raw_TOWNSHIP == "Commercial Twp" |raw_TOWNSHIP == "Deerfield Twp"  | raw_TOWNSHIP == "Fairfield Twp")

capemay_county_data <- nj_statewide_data %>%
  filter(raw_TOWNSHIP == "Lower Twp" |raw_TOWNSHIP == "Middle Twp")

salem_county_data <- nj_statewide_data %>%
  filter(raw_TOWNSHIP == "Pennsville Twp" |raw_TOWNSHIP == "Pittsgrove Twp"  | raw_TOWNSHIP == "Carneys Point Twp")


```

Manipulating philly tibble to only look at black and white races

```{r}
philly_data %>% 
  select(raw_row_number, date, subject_race, raw_race, arrest_made, search_conducted) -> philly_data

philly_data %>% 
  filter(subject_race == "white" | subject_race == "black") %>% 
  filter(date >= '2016-01-01' & date < '2017-01-01') %>%   
  mutate(arrest_made = as.numeric(arrest_made)) %>% 
  mutate(search_conducted = as.numeric(search_conducted)) -> philly_data

philly_white_data <- philly_data %>% 
  filter(subject_race == "white")

philly_black_data <- philly_data %>%
  filter(subject_race == "black")

```


Finding prop of each race searched for every county

```{r}
#unique(gloucester_county_data$search_conducted)

# Assume if search_conducted is NA, then there is no search

gloucester_county_data %>%
  filter(raw_RACE == "W - WHITE" | raw_RACE == "B - BLACK") %>%
  mutate(search_conducted = if_else(is.na(search_conducted) | search_conducted == FALSE, 0, 1)) %>%
  group_by(raw_RACE) %>%
  summarize(mean(search_conducted)) -> gloucester_search_prop

cumberland_county_data %>%
    filter(raw_RACE == "W - WHITE" | raw_RACE == "B - BLACK") %>%
  mutate(search_conducted = if_else(is.na(search_conducted)| search_conducted == FALSE, 0, 1)) %>%
  group_by(raw_RACE) %>%
  summarize(mean(search_conducted)) -> cumberland_search_prop

capemay_county_data %>%
    filter(raw_RACE == "W - WHITE" | raw_RACE == "B - BLACK") %>%
  mutate(search_conducted = if_else(is.na(search_conducted)| search_conducted == FALSE, 0, 1)) %>%
  group_by(raw_RACE) %>%
  summarize(mean(search_conducted)) -> capemay_search_prop

salem_county_data %>%
    filter(raw_RACE == "W - WHITE" | raw_RACE == "B - BLACK") %>%
  mutate(search_conducted = if_else(is.na(search_conducted)| search_conducted == FALSE, 0, 1)) %>%
  group_by(raw_RACE) %>%
  summarize(mean(search_conducted)) -> salem_search_prop

camden_data %>%
    filter(raw_RACE == "W - WHITE" | raw_RACE == "B - BLACK") %>%
  mutate(search_conducted = if_else(is.na(search_conducted)| search_conducted == FALSE, 0, 1)) %>%
  group_by(raw_RACE) %>%
  summarize(mean(search_conducted)) -> camden_search_prop


```







Finding proportion of searches for black and white in Philly

```{r}
philly_white_data %>%
  summarize(white_search_prop = mean(search_conducted))

# 0.05522

philly_black_data %>%
  summarize(black_search_prop = mean(search_conducted))
# 0.06309
```


Making confidence interval

```{r}
camden_data %>%
  rep_sample_n(reps = 1000, replace = T, size = 50) -> camden_boot_samp

philly_black_data %>%
   rep_sample_n(reps = 1000, replace = T, size = 50) -> philly_black_boot

black_boot_props <- philly_black_boot %>%
  group_by(replicate) %>%
  summarize(black_search_prop = mean(search_conducted))


```

Plotting Bootstrap Sample

```{r}
ggplot(black_boot_props, aes(x = black_search_prop)) +
  geom_histogram(fill = "darkseagreen3", bins = 10, color = "white") + 
  xlab("Black Search Proportions in Bootstrap Samples") + 
  ylab("Frequency") + 
  theme_bw()

```

Confidence Interval

```{r}
quantile(black_boot_props$black_search_prop, c(0.05, 0.95) )

```

New Jersey Stuff

```{r}
nj_counties <- rbind(gloucester_county_data, cumberland_county_data, capemay_county_data, salem_county_data)

# Filtering by only white and black race, and data from 2016
nj_counties %>% 
 filter(raw_RACE == "W - WHITE" | raw_RACE == "B - BLACK") %>%
  filter(date >= '2016-01-01' & date < '2017-01-01') %>%   
  mutate(search_conducted = if_else(is.na(search_conducted)| search_conducted == FALSE, 0, 1)) %>%
  group_by(raw_RACE) %>%
  summarize(mean(search_conducted)) -> nj_counties_search_prop
  
```

```{r}


g <- ggplot(nj_counties_search_prop, aes(x = raw_RACE, y = `mean(search_conducted)`, fill = raw_RACE)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Search Rate by Race For NJ Counties (2016)",
       x = "Race",
       y = "Search Rate") +
  theme_minimal() +
  scale_y_continuous(breaks = seq(0, max(nj_counties_search_prop$`mean(search_conducted)`), by = 0.01)) + # Add more y-axis labels
  scale_fill_manual(values = c("White" = "blue", "Black" = "green"))

ggsave("nj_counties_search_rate.png", plot = g )


```
```{r}
ggplot(camden_search_prop, aes(x = raw_RACE, y = `mean(search_conducted)`, fill = raw_RACE)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Search Rate by Race For NJ Counties (2016)",
       x = "Race",
       y = "Search Rate") +
  theme_minimal() -> g2

ggsave("camden_search_prop.png", plot = g2 )
```
BOOTSTRAPS FOR NJ COUNTIES

```{r}
nj_counties <- nj_counties %>% 
 filter(raw_RACE == "W - WHITE" | raw_RACE == "B - BLACK") %>%
  filter(date >= '2016-01-01' & date < '2017-01-01') %>%   
  mutate(search_conducted = if_else(is.na(search_conducted)| search_conducted == FALSE, 0, 1)) 

nj_counties_black <- nj_counties %>%
  filter(raw_RACE == "B - BLACK")

nj_counties_white <- nj_counties %>%
  filter(raw_RACE == "W - WHITE")



```

Test stats for black and white in CAMDEN

```{r}
nj_counties_black %>%
  summarize(black_search_prop = mean(search_conducted))

# 0.0733

nj_counties_white %>%
  summarize(white_search_prop = mean(search_conducted))
# 0.0288

```

Bootstraps

```{r}
nj_counties_black %>%
  rep_sample_n(size = nrow(nj_counties_black), replace = TRUE, reps = 1000) -> bootstrap_samples

black_boot_props <- bootstrap_samples %>%
  group_by(replicate) %>%
  summarize(black_search_prop = mean(search_conducted))

ggplot(black_boot_props, aes(x = black_search_prop)) +
  geom_histogram(fill = "darkseagreen3", bins = 10, color = "white") + 
  xlab("Black Search Proportions in Bootstrap Samples") + 
  ylab("Frequency") + 
  theme_bw() -> g3
g3

ggsave("njcounties_bootstrap_distrib.png", plot = g3)

quantile(black_boot_props$black_search_prop, c(0.05, 0.95) )

```

```{r}
nj_counties_white %>%
  rep_sample_n(size = nrow(nj_counties_white), replace = TRUE, reps = 1000) -> bootstrap_samples

white_boot_props <- bootstrap_samples %>%
  group_by(replicate) %>%
  summarize(white_search_prop = mean(search_conducted))

ggplot(white_boot_props, aes(x = white_search_prop)) +
  geom_histogram(fill = "darkseagreen3", bins = 10, color = "white") + 
  xlab("White Search Proportions in Bootstrap Samples") + 
  ylab("Frequency") + 
  theme_bw() -> g4

g4
ggsave("njcounties_bootstrap_distrib_WHITE.png", plot = g4)


quantile(white_boot_props$white_search_prop, c(0.05, 0.95) )

```

BOOTSTRAP FOR CAMDEN

```{r}
camden_data %>% 
  filter(raw_RACE == "W - WHITE" | raw_RACE == "B - BLACK") %>%
  filter(date >= '2016-01-01' & date < '2017-01-01') %>%   
  mutate(search_conducted = if_else(is.na(search_conducted)| search_conducted == FALSE, 0, 1)) -> camden_data


camden_data %>%
  filter(raw_RACE == "B - BLACK") -> camden_black

camden_data %>%
  filter(raw_RACE == "W - WHITE") -> camden_white

camden_black %>%
  rep_sample_n(size = nrow(camden_black), replace = TRUE, reps = 1000) -> camden_black_bootstrap

camden_black_boot_props <- camden_black_bootstrap %>%
  group_by(replicate) %>%
  summarize(black_search_prop = mean(search_conducted))

ggplot(camden_black_boot_props, aes(x = black_search_prop)) +
  geom_histogram(fill = "darkseagreen3", bins = 10, color = "white") + 
  xlab("Black Search Proportions in Bootstrap Samples") + 
  ylab("Frequency") + 
  theme_bw() -> camden_black_boot_graph
camden_black_boot_graph

ggsave("camden_black_bootstrap.png", camden_black_boot_graph)

quantile(camden_black_boot_props$black_search_prop, c(0.05, 0.95) )



```

CAMDEN WHITE BOOTSTRAP

```{r}

camden_white %>%
  rep_sample_n(size = nrow(camden_white), replace = TRUE, reps = 1000) -> camden_white_bootstrap

camden_white_boot_props <- camden_white_bootstrap %>%
  group_by(replicate) %>%
  summarize(white_search_prop = mean(search_conducted))

ggplot(camden_white_boot_props, aes(x = white_search_prop)) +
  geom_histogram(fill = "darkseagreen3", bins = 10, color = "white") + 
  xlab("White Search Proportions in Bootstrap Samples") + 
  ylab("Frequency") + 
  theme_bw() -> camden_white_boot_graph
camden_white_boot_graph

ggsave("camden_white_bootstrap.png", camden_white_boot_graph)

quantile(camden_white_boot_props$white_search_prop, c(0.05, 0.95) )

```

Philly hypothesis test

```{r}
# Philly is already filtered for only black and white

philly_data %>%
  group_by(subject_race) %>%
  summarize(prop_searched = mean(search_conducted)) -> observed_props

# Extract observed proportion for Black
observed_black_prop <- observed_props %>%
  filter(subject_race == "black") %>%
  pull(prop_searched)

# Extract observed proportion for White
observed_white_prop <- observed_props %>%
  filter(subject_race == "white") %>%
  pull(prop_searched)

observed_diff <- observed_black_prop - observed_white_prop

# Null distribution for difference in proportions (assuming no race effect)
null_distribution <- philly_data %>%
  specify(response = search_conducted) %>%
  hypothesize(null = "paired independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "mean")


```



